---
title: "Clinical Trials 101 Course"
---

## Study Objectives

### 1.1 Primary Objectives

#### 1.1.1 Objectives

(1)  To evaluate the efficacy of intravenous ketorolac compared to placebo in the treatment of acute migraine in combination with intravenous 0.9% normal saline and metoclopramide as measured by the **mean reduction in pain intensity on the Visual Analogue Scale (VAS)** in children or adolescents between 5 and 17 years of age who present to the ED and have not achieved a satisfactory response from usual home therapy or at least one oral analgesic.

(2)  To evaluate the **safety and tolerability** of intravenous ketorolac and metoclopramide in the treatment of acute migraine in children or adolescents between 5 and 17 years of age who present to the ED and have not achieved a satisfactory response from usual home therapy or at least one oral analgesic.

#### 1.1.2 Hypothesis

(1)  Intravenous ketorolac administered in combination with metoclopramide and 0.9% normal saline is superior to intravenous metoclopramide and 0.9% normal saline alone as measured by the mean reduction in pain intensity on the Visual Analogue Scale (VAS) at 2 hours in children 5 to 17 years of age who present to the ED and have not achieved a satisfactory response from usual home therapy or at least one oral analgesic.

### 1.2 Secondary Objectives

#### 1.2.1 Objectives 

(1)  To evaluate the efficacy of intravenous ketorolac compared to placebo in the treatment of acute migraine in combination with intravenous 0.9% normal saline and metoclopramide as measured by **pain freedom** **(VAS=0) at 2 hours** in children or adolescents between 5 and 17 years of age who present to the ED and have not achieved a satisfactory response from usual home therapy or at least one oral analgesic.

(2)  To evaluate the efficacy of intravenous ketorolac compared to placebo in the treatment of acute migraine in combination with intravenous 0.9% normal saline and metoclopramide as measured by **headache relief defined as a 33% reduction on the VAS at 2 hours** in children or adolescents between 5 and 17 years of age who present to the ED and have not achieved a satisfactory response from usual home therapy or at least one oral analgesic.

(3)  To evaluate the efficacy of intravenous ketorolac compared to placebo in the treatment of acute migraine in combination with intravenous 0.9% normal saline and metoclopramide as measured by **headache relief defined as a 50% reduction on the VAS at 2 hours** in children or adolescents between 5 and 17 years of age who present to the ED and have not achieved a satisfactory response from usual home therapy or at least one oral analgesic.

#### 1.2.2 Hypothesis

(1)  Intravenous ketorolac administered in combination with metoclopramide and 0.9% normal saline is superior to intravenous metoclopramide and 0.9% normal saline alone as measured by the secondary outcome measures described above in children 5 to 17 years of age who present to the ED and have not achieved a satisfactory response from usual home therapy or at least one oral analgesic.

### 1.3 Tertiary objectives

(1)  To evaluate the efficacy of intravenous ketorolac compared to placebo in the treatment of acute migraine in combination with intravenous 0.9% normal saline and metoclopramide as measured by **sustained pain-freedom (no recurrence of headache within 24 hours** if pain was completely eliminated (VAS = 0) prior to discharge) in children or adolescents between 5 and 17 years of age who present to the ED and have not achieved a satisfactory response from usual home therapy or at least one oral analgesic.

\(1\) To evaluate the efficacy of intravenous ketorolac compared to placebo in the treatment of acute migraine in combination with intravenous 0.9% normal saline and metoclopramide as measured by **sustained headache relief (no increase in headache by 33% on the VAS or 50% on the VAS if headache relief was initially achieved)** in children or adolescents between 5 and 17 years of age who present to the ED and have not achieved a satisfactory response from usual home therapy or at least one oral analgesic.

(2) To explore the concept of a **minimum clinically significant difference** at 2 hours correlating the VAS and patient-centric outcomes including (1) \"I would take the medication again\" (19) (2) \"my headache is a bit better/worse\" and \"my headache is a lot better/worse\" (20), and (3) the treating physicians subjective assessment of efficacy all assessed at the 2 hour time point.

(3) To compare the VAS and Faces Pain Scale -- Revised (FPS-R) in children \< 8 years.

\(4\) A descriptive economic analysis of ED migraine treatment.

## Results

An analysis of a randomized controlled trial of ketorolac + metoclopramide versus metoclopramide alone.

```{r}
#| echo: false
#| output: false
#| warning: false

library(rsconnect)
library(tidyverse)
library(consort)
library(skimr)
source("keto_migraine_rct.R")

```

```{r}
#| echo: false
#| label: consort-plot
#| fig-cap: "CONSORT Diagram"
set.seed(1001)
N <- 645

sid <- 1:N
randomized <- rep("0", N)
randomized[sample(1:N,54)] <- "1"

exc1 <- rep(NA, N)
exc1[randomized=="0"] <- sample(c("Not meeting inclusion criteria", "Declined to participate",
                                  "Other reasons"), 591, replace = TRUE, prob = c(0.9, 0.06, 0.04))

exc2 <- rep(NA, N)

# randomized[is.na(exc1)] <- "1"

df <- tibble(sid,randomized,exc1)

df_r <- df %>%
  filter(randomized=="1") %>%
  mutate(id=row_number())

out <- consort_plot(data = df,
                    orders = c(sid = "Assessed for eligibility",
                               exc1    = "Excluded",
                               randomized    = "Randomized"),
                    side_box = c("exc1"),
                    labels = c("1" = "Screening", "2" = "Randomization"))

plot(out)


```
